### 2022-10-30 dissertation/chapters/05-probability.tex
✨ (chapter5): Review Done

      \item \textbf{Section \ref{sec:probability:probability_distributions}} presents the reader with the details of some relevant probability distributions.

@@

      \item \textbf{Section \ref{sec:probability:bayesian_statistics}} presents \textit{Bayesian} statistics. Brief discussions follow on the \textit{frequentist} view and \textit{Bayesian} view of probability. Detailed discussions follow on \index{Bayesian optimisation}Bayesian optimisation methods such as \index{Bayesian analysis}\textit{Bayesian analysis}.

@@

In everyday conversation, the term \textit{probability} is a measure of one's belief in the occurrence of a future event~\cite{ref:wackerly:2014}. Probability is a necessary tool used in many fields including physics, biology, chemistry and computer science. These fields contain many cases that generate observations that cannot be predicted with absolute certainty~\cite{ref:wackerly:2014}. Probability can be be inferred and confirmed through past events. These events are referred to as \textit{random} or \textit{stochastic} events. The probability that a certain event $A$ might occur is denoted $P(A)$. Although these random events cannot be predicted with absolute certainty, the relative frequency with which they occur over many trials, is often remarkably stable.

Consider flipping an unbiased, fair coin. The coin has two possible outcomes. From this, one can conclude that each side has a $\frac{1}{2}$ or $50\%$ chance of occurring. In statistics, the decimal probability notation is used, where $0 <= P(A) <= 1$. Suppose the fair coin is thrown 10 times, one is not guaranteed to observe $0.5$ heads and $0.5$ tails. There is some probability (although small) that the coin might fall on heads $0/10$ times. The probability of such an event occurring is $0.0009765625$. In the coin flip example, the \acf{CLT} shows that the normalised sum of events tends toward a normal distribution with a mean value of $0.5$ if the number of events observed $N$, is large~\cite{ref:wackerly:2014}. The larger the value of $N$, the higher the confidence of mean probability and relative frequency of the event. The stable long-term relative frequency by which a random event occurs, provides an intuitive and meaningful measure of belief that a certain event will occur again at some point in the future~\cite{ref:wackerly:2014}.

@@

The occurrence of a given random event $A$ can often be conditional on the occurrence of another event $B$. In the field of medicine, an example of this is to calculate the probability of a certain diagnosis of a sick patient given his/her symptoms. The aforementioned relationship between events is referred to as the conditional probability between two events. The conditional probability is expressed as $P(A \vert B)$ and is read \textit{the probability of $A$ given $B$}. On the contrary, the \textit{unconditional} probability is the probability of an event, not dependent on any other. The conditional probability of $A$ given $B$ can be expressed as is given in Definition \ref{eq:probability:cond_probability} below~\cite{ref:wackerly:2014}.

\begin{definition}[\textbf{Conditional Probability}]
      \label{eq:probability:cond_probability}
      The conditional probability of an event $A$, subject to the occurrence of event $B$ is expressed as

@@

It could also be the case that the outcome of observing event $A$ is not affected by the occurrence of $B$. In this case, it is said that events $A$ and $B$ are independent. The independence between two events are expressed in Definition \ref{def:probability:cond_probability:independence} below.

\begin{definition}[\textbf{Independence of Events}]
      \label{def:probability:cond_probability:independence}
      Two events, $A$ and $B$ are said to be independent of each other if, and only if the following criteria hold:

@@

      \noindent
      otherwise events $A$ and $B$ are said to be \textit{dependent} random events.
\end{definition}

@@

Suppose there are two random events $A$ and $B$, then one can calculate the probability of the union and intersection of these events. From the aforementioned concept, two laws of probability can be formulated. These are referred to as the \textit{multiplicative} and \textit{additive} laws of probability~\cite{ref:wackerly:2014} and is given below in Theorems \ref{th:probability:multiple_events:multiplicative} and \ref{th:probability:multiple_events:additive} respectively.

@@

\begin{proof}
      Proof is given from Definition \ref{eq:probability:cond_probability} above.
\end{proof}

@@

Bayes' Theorem, named after Thomas Bayes, describes the probability of an event $A$, based on prior knowledge of conditions that might be related to $A$~\cite{ref:zalta:2015}. In order to derive the formal theorem and proof, first consider Definition \ref{def:probability:bayes_theorem:partition} below~\cite{ref:zalta:2015}.

\begin{definition}
      \label{def:probability:bayes_theorem:partition}
      For some positive integer, $K$, let the sets $B_{1}, B_{2}, \dots, B_{K}$ be such that

@@

      \noindent
      Then the collection of sets $\{B_{1}$, $B_{2}$, $\dots$, $B_{K}\}$ is said to be a partition of S, the union of mutually exclusive subsets.
\end{definition}

@@

\begin{proof}
      The proof follows from the definition of conditional probability as was presented in \ref{sec:probability:cond_probability}.

@@

The \index{Dirichlet probability distribution}Dirichlet probability distribution is a family of multivariate continuous probability distributions over some $x$ in $K$ dimensions~\cite{ref:wackerly:2014}. The \index{Dirichlet probability distribution}Dirichlet probability distribution is the multivariate generalization of the \index{Beta probability distribution}Beta probability distribution and is thus sometimes referred to by its alternative name, the multivariate Beta probability distribution. The \index{Dirichlet probability distribution}Dirichlet probability distribution is parameterised by some vector $\vec{\alpha} = (\alpha_{1}$,  $\dots, \alpha_{K})$ $\forall_{k=1}^{K} \alpha_{k} > 0$, $\alpha_{k} \in \mathbb{R}$. $\alpha$ is referred to as the concentration parameter. The \index{Dirichlet probability distribution}Dirichlet probability distribution of order $K \geq 2$ with parameters $\alpha$, denoted $Dir(\alpha)$, has a \acs{PDF} as is given in Equation~\eqref{eq:probability:probability_distributions:dirichlet:pdf} below.

@@

\noindent
while the mode of the distribution is given in Equation~\eqref{eq:probability:probability_distributions:bernoulli:mode} below.

@@

\subsection{Binomial Distribution}\label{sec:probability:probability_distributions:bin}

The \index{Binomial distribution}Binomial distribution is a discrete probability distribution over a random variable $x$ taking on a number of successes in $N$ sequential independent experiments that each ask a \textit{yes-no} question~\cite{ref:wackerly:2014}. The probability of a single independent experiment yielding a success, is given as $\theta$ and the \index{Binomial distribution}Binomial distribution is denoted as $Bin(N, \theta)$ with support $x \in \{0, 1, \dots, N\}$.  It should be noted that the \index{Binomial distribution}Binomial distribution is the extension of the \index{Bernoulli probability distribution}Bernoulli probability distribution over $N$ independent sequential experiments and thus, each experiment also yields some boolean outcome. When $N=1$, the experiment is referred to as a Bernoulli trial and the distribution is just a \index{Bernoulli probability distribution}Bernoulli probability distribution. When $N > 1$, the sequence of outcomes is referred to as a Bernoulli process. The \acf{PMF} of the \index{Binomial distribution}Binomial distribution is given in Equation~\eqref{eq:probability:probability_distributions:binomial:pmf} below.

@@

\noindent
Similar to the \index{Bernoulli probability distribution}Bernoulli probability distribution, the mean of the \index{Binomial distribution}Binomial distribution is $N\theta$ given the \acs{CLT} as was shown in Figure \ref{fig:probability:probability_distributions:bernoulli:coin}. The expected value of the \index{Binomial distribution}Binomial distribution is thus given in Equation~\eqref{eq:probability:probability_distributions:binomial:expected_value} below.

@@

The \index{Categorical probability distribution}Categorical probability distribution is a discrete probability distribution over some random variable $x$, taking on any one of $K$ possible categories~\cite{ref:wackerly:2014}. There is no ordering to these categories and therefore, for simplicity, each category is assigned a numerical representative value such that $k = (1, 2, \dots, K)$. The probabilities for all outcomes is given by the probability vector $\vec{\theta} = (\theta_{1}, \dots, \theta_{K})$.  This means that the probability $P(x=k)=\theta_{k}$, has support $x \in \{1, \dots, K\}$. The \index{Categorical probability distribution}Categorical probability distribution, denoted $Cat(\theta)$, is the generalization of the \index{Bernoulli probability distribution}Bernoulli probability distribution and is sometimes referred to it by its alternative names, the generalised \index{Bernoulli probability distribution}Bernoulli probability distribution or the Multinoulli probability distribution. In probability theory, the \index{Categorical probability distribution}Categorical probability distribution is often used to explain the outcome of a single experiment with more than two possible outcomes such as rolling a die~\cite{ref:wackerly:2014}. The \acs{PMF} of the \index{Categorical probability distribution}Categorical probability distribution is given in Equation~\eqref{eq:probability:probability_distributions:categorical:pmf} below.

@@

The \index{Multinomial probability distribution}Multinomial probability distribution is a discrete probability distribution over some random variable $x = (x_{1}, \dots\, x_{K})$ that takes on the counts for each occurrence of $K$ possible classes in $N$ independent trials~\cite{ref:wackerly:2014}. The probabilities for all possible outcomes in a single trial is given by the probability vector $\vec{\theta} = (\theta_{1}, \dots, \theta_{K})$. The \index{Multinomial probability distribution}Multinomial probability distribution, denoted $Mul(N, K, \theta)$, is thus the generalization of the \index{Binomial distribution}Binomial distribution to $K$ dimensions. Consider the following special cases.

\begin{itemize}
      \item When $K$ is 2 and $N = 1$, the \index{Multinomial probability distribution}Multinomial probability distribution is the \index{Bernoulli probability distribution}Bernoulli probability distribution.
      \item When $K$ is 2 and $N > 1$, the \index{Multinomial probability distribution}Multinomial probability distribution is the \index{Binomial distribution}Binomial distribution.
      \item When $K > 2$ and $N = 1$, the \index{Multinomial probability distribution}Multinomial probability distribution is the \index{Categorical probability distribution}Categorical probability distribution.
\end{itemize}

@@

\noindent
Similar to the \index{Categorical probability distribution}Categorical probability distribution, the random variable $x$ can also be encoded in binary format, yielding an $N \times K$ matrix $X$ of \index{Bernoulli probability distribution}Bernoulli probability distributions. The support is then given as $X \in \{0, 1\}^{N \times K}, \forall_{i=1}^{N}\sum_{k=1}^{K} x_{i,k} = 1$ so that the standard $K-1$ probability simplex $S$ still holds for each trial. The \acs{PMF} of the \index{Multinomial probability distribution}Multinomial probability distribution can then be rewritten as is presented in Equation~\eqref{eq:probability:probability_distributions:multinomial:pmf_alt}

@@

\noindent
In Equation~\eqref{eq:probability:probability_distributions:multinomial:pmf_alt} above, $N_{k}$ is a summary variable denoting the number of times a category $k$ occurs over all trials in $N$.

@@

\citeauthor{ref:wackerly:2014}\cite{ref:wackerly:2014} states that conjugate priors are prior probability distributions that results in posterior distributions that are of the same functional form $\mathcal{A}(v)$ as the prior, but with different parameter values. This section considers the conjugate priors for the Binomial likelihood and Categorical/Multinomial likelihood.

@@

      \item Let $X = (x_{1}, \dots, x_{I})$ be the outcomes of independent, identical random events, each with boolean outcome. That is $x_{i} | \theta \overset{\text{iid}}{\sim} Ber(\theta)$ and thus $\mathcal{L}(x_{i} \vert \theta)$ is the Bernoulli likelihood.

@@

\noindent
The Bernoulli likelihood is then given in Equation~\eqref{eq:probability:conjugate_priors:binom_likelihood:likelihood} below.

@@

\noindent
The posterior distribution has the same functional form $\mathcal{A}(v)$ as the prior, but with updated prior parameters $\alpha' = N_{1} + \alpha$ and $\beta' = N_{0} + \beta$. This shows that the \index{Beta probability distribution}Beta probability distribution is the conjugate prior to the Bernoulli likelihood.

@@

The conjugate prior to a \index{Categorical probability distribution}Categorical and \index{Multinomial probability distribution}Multinomial probability distribution is the \index{Dirichlet probability distribution}Dirichlet probability distribution~\cite{ref:wackerly:2014}. The proof of the conjugate prior for the \index{Bernoulli probability distribution}Bernoulli probability distribution is similar to the proof presented in Section \ref{sec:probability:conjugate_priors:binom_likelihood} above.This means that the posterior distribution must have the same functional form $\mathcal{A}(v)$ as the prior distribution. The proof is shown as follows. \\\\
\textbf{Setup}:

@@

      \item Let $X = (x_{1}, \dots, x_{I})$ be the outcomes of independent, identical random events, each with $K$ possible outcomes. That is $x_{i} | \theta \overset{\text{iid}}{\sim} Cat(\theta)$ and thus $\mathcal{L}(x_{i} \vert \theta)$ is the Categorical likelihood.

@@

\noindent
The likelihood of the Categorical and \index{Multinomial probability distribution}Multinomial probability distributions is then given in Equation~\eqref{eq:probability:conjugate_priors:mult_likelihood:likelihood} below.

@@

\noindent
The posterior distribution has the same form $\mathcal{A}(v)$ as the prior, but with updated prior parameters $\alpha' = N + \alpha$. This shows that the \index{Dirichlet probability distribution}Dirichlet probability distribution is the conjugate prior to the Categorical and Multinomial likelihood.

@@

In general, there are two main views to probability and statistics. These include the \textit{frequentist} and the \textit{Bayesian} view of statistics. Naturally, Bayesian statistics is based on Bayes' Theorem as was presented in Section \ref{sec:probability:bayes_theorem}. Bayesian statistics describe the probability of an event in terms of some belief based on previous knowledge of the event and the conditions under which the event happens~\cite{ref:hackenberger:2019}. To introduce the concept of Bayesian inference and analysis, one first need to understand the difference between these two approaches. Bayesian statistics out-date the frequentist approach, but lacked interest in the early days partly because of the limited applications where the conjugate priors where known~\cite{ref:hackenberger:2019}. More recent advancements in mathematical methods popularised the Bayesian approach again. A notable contribution to this switch was the invention of \acs{MCMC} in the 1950's. This family of algorithms allowed for the construction of random sampling algorithms from a probability distribution which allows for the calculation of Bayesian hierarchical models~\cite{ref:hackenberger:2019}. Soon after followed one of the earliest papers that use Bayesian statistics in the field of medicine in 1982~\cite{ref:ashby:2006}.

The difference between the frequentist approach and the Bayesian approach can be illustrated using an example.~\citeauthor{ref:hackenberger:2019}~\cite{ref:hackenberger:2019} suggests an experiment that investigates whether the sex ration in some hypothetical mice population is $1:1$. Two experiments can be designed. In the first experiment, mice are randomly selected until the first male is chosen. The result in this experiment will then be the total number of mice chosen by sex. For the second experiment, exactly seven mice are randomly selected. The result of the second experiment would be the number of males and females in a sample of seven. Suppose the outcome of the experiment was $FFFFFFM$, where $F$ represents a female and $M$ represents a male. If the experimental design is not known ahead of time, the result is useless. Consider the $P$-value for each of these experiments. The P-value is the probability of obtaining results at least as extreme as the observed results of a statistical hypothesis test~\cite{ref:beers:2022}. For the first experiment, the $P$-value is $0.031$ and for the second experiment, the $P$-value is $0.227$. Using a confidence level of $\alpha = 0.05$, one would conclude opposite outcomes for these two experiments when it comes to rejecting the null hypothesis, despite using the same data. The reason for this difference is because of the difference in their null distributions. The first approach uses a geometrical approach and the second used a binomial approach as is illustrated on Figure \ref{fig:probability:bayesian_statistics:mouse_experiment_outcome} below.

\begin{figure}[htbp]
      \includegraphics[width=\textwidth]{images/mouse_experiment_outcome.jpg}
      \caption{The experimental outcomes for the mice-population experiments as was taken from~\cite{ref:hackenberger:2019} .}
      \label{fig:probability:bayesian_statistics:mouse_experiment_outcome}
\end{figure}

If Bayesian statistics is used then the experimental design that was chosen does not matter. In Bayesian statistics it is common to use a \index{Beta probability distribution}Beta probability distribution as a prior distribution. If the prior distribution is sampled from $Beta(3,3)$ then using Bayesian analysis, the posterior distribution, according to the outcomes of this experiment, would yield $Beta(9,4)$. The process by which this is done is shown later in this section. The $Beta$ distribution was presented in Section \ref{sec:probability:probability_distributions:beta}.~\citeauthor{ref:hackenberger:2019}~\cite{ref:hackenberger:2019} mentions that the $Beta$ distribution can be seen as a probability distribution of the occurrence of specific parameters. From the information that is now known about the $Beta$ distribution, it is possible to calculate the probability that the sex ratio in this mice population is not 1:1 with the $Beta$ distribution as a prior, yielding a $P$-value of $0.92$. This means that there is a probability of $92\%$ that the mice population is not 1:1, regardless of experiment design. This is illustrated in Figure \ref{fig:probability:bayesian_statistics:mouse_distributions} below.

\begin{figure}[htbp]
      \includegraphics[width=\textwidth]{images/mouse_experiment_distributions.jpg}
      \caption{An illustration of the prior and posterior probability distributions for the outcomes of the mice-population experiment, using a $Beta$ prior, as was taken from~\cite{ref:hackenberger:2019}.}
      \label{fig:probability:bayesian_statistics:mouse_distributions}
\end{figure}

One can see from Figure \ref{fig:probability:bayesian_statistics:mouse_distributions} how the posterior distribution differs from the prior distribution. This is the nature of Bayesian analysis and is presented in more detail next.

@@

The marginal density or mass function of $Y_{1}, Y_{2}, \dots, Y_{N}$ is given in Equation~\eqref{eq:probability:bayesian_statistic:bayesian_analysis:marginal_density} below.

@@

Finally, the posterior density of $\theta \vert y_{1}, y_{2}, \dots, y_{n}$ according to Bayes' Theorem is given in Equation~\eqref{eq:probability:bayesian_statistic:bayesian_analysis:posterior_density} below.

>>>>

      \item \textbf{Section \ref{sec:probability:probability_distributions}} presents the reader with relevant probability distributions.

@@

      \item \textbf{Section \ref{sec:probability:bayesian_statistics}} presents \textit{Bayesian} statistics. Brief discussions follow on the \textit{frequentist} view and \textit{Bayesian} view of probability. Detailed discussions follow on Bayesian optimisation methods such as \index{Bayesian analysis}\textit{Bayesian analysis}.

@@

In everyday conversation, the term \textit{probability} is a measure of one's belief in the occurrence of a future event~\cite{ref:wackerly:2014}. Probability is a necessary tool used in many fields including physics, biology, chemistry and computer science. These fields contain many cases that generate observations that cannot be predicted with absolute certainty~\cite{ref:wackerly:2014}. Probability can be be inferred and confirmed through past events. These events are referred to as \textit{random} or \textit{stochastic} events. The probability that a certain event $A$ might occur, is denoted $P(A)$. Although these random events cannot be predicted with absolute certainty, the relative frequency with which they occur over many trials, is often remarkably stable.

Consider flipping an unbiased, fair coin. The coin has two possible outcomes. It can conclude that each side has a $\frac{1}{2}$ or $50\%$ chance of occurring. In statistics, the decimal probability notation is used, where $0 <= P(A) <= 1$. Suppose the fair coin is thrown 10 times, one is not guaranteed to observe $0.5$ heads and $0.5$ tails. There is some probability (although small) that the coin might fall on heads $0/10$ times. The probability of such an event occurring is $0.0009765625$. In the coin flip example, the \acf{CLT} shows that the normalised sum of events tends toward a normal distribution with a mean value of $0.5$ if the number of events observed $N$, is large~\cite{ref:wackerly:2014}. The larger the value of $N$, the higher the confidence of mean probability and relative frequency of the event. The stable long-term relative frequency by which a random event occurs, provides an intuitive and meaningful measure of belief that a certain event will occur again at some point in the future~\cite{ref:wackerly:2014}.

@@

The occurrence of a given random event $A$ can often be conditional on the occurrence of another event $B$. In the field of medicine, an example of this is to calculate the probability of a certain diagnosis of a sick patient given his/her symptoms. The aforementioned relationship between events is referred to as the conditional probability between two events. The conditional probability is expressed as $P(A \vert B)$ and is read \textit{the probability of $A$ given $B$}. On the contrary, the \textit{unconditional} probability is the probability of an event, not dependent on any other. The conditional probability of $A$ given $B$ can be expressed as is given in Definition \ref{eq:probability:cond_probability} below~\cite{ref:wackerly:2014}.
\\
\begin{definition}[\textbf{Conditional Probability}]
      \label{eq:probability:cond_probability}
      The conditional probability of an event $A$, subject to the occurrence of event $B$ is expressed as

@@

It could also be the case that the outcome of observing event $A$ is not affected by the occurrence of $B$. In this case, it is said that events $A$ and $B$ are independent. The independence between two events are expressed in Definition \ref{def:probability:cond_probability:independence} below.
\\
\begin{definition}[\textbf{Independence of Events}]
      \label{def:probability:cond_probability:independence}
      Two events, $A$ and $B$ are said to be independent of each other if, and only if the following criteria hold:

@@

      \noindent
      otherwise, events $A$ and $B$ are said to be \textit{dependent} random events.
\end{definition}

@@

Suppose there are two random events $A$ and $B$, then one can calculate the probability of the union and intersection of these events. From the aforementioned concept, two laws of probability can be formulated. These are referred to as the \textit{multiplicative} and \textit{additive} laws of probability~\cite{ref:wackerly:2014} and is given below in Theorems \ref{th:probability:multiple_events:multiplicative} and \ref{th:probability:multiple_events:additive} respectively.\\

@@

\begin{proof}
      Proof is given from Definition \ref{eq:probability:cond_probability} above.
\end{proof}
\vspace*{0.5cm}

@@

Bayes' Theorem, named after Thomas Bayes, describes the probability of an event $A$, based on prior knowledge of conditions that might be related to $A$~\cite{ref:zalta:2015}. In order to derive the formal theorem and proof, first consider Definition \ref{def:probability:bayes_theorem:partition} below~\cite{ref:zalta:2015}.
\\
\begin{definition}
      \label{def:probability:bayes_theorem:partition}
      For some positive integer, $K$, let the sets $B_{1}, B_{2}, \dots, B_{K}$ be such that

@@

      \noindent
      Then the collection of sets $\{B_{1}$, $B_{2}$, $\dots$, $B_{K}\}$ is said to be a partition of S, the union of mutually exclusive subsets.
\end{definition}
\vspace*{0.5cm}

@@

\begin{proof}
      The proof follows from the definition of conditional probability as was presented in Section~\ref{sec:probability:cond_probability}.

@@

The \index{Dirichlet probability distribution}Dirichlet probability distribution is a family of multivariate continuous probability distributions over some $x$ in $K$ dimensions~\cite{ref:wackerly:2014}. The \index{Dirichlet probability distribution}Dirichlet probability distribution is the multivariate generalization of the \index{Beta probability distribution}Beta probability distribution and is thus sometimes referred to by its alternative name, the multivariate Beta probability distribution. The \index{Dirichlet probability distribution}Dirichlet probability distribution is parameterised by some vector $\vec{\alpha} = (\alpha_{1}$,  $\dots, \alpha_{K})$ $\forall_{k=1}^{K} \alpha_{k} > 0$, $\alpha_{k} \in \mathbb{R}$. Throughout this chapter, the vector notation is dropped for simplicity and readability. $\alpha$ is referred to as the concentration parameter. The \index{Dirichlet probability distribution}Dirichlet probability distribution of order $K \geq 2$ with parameters $\alpha$, denoted $Dir(\alpha)$, has a \acs{PDF} as is given in Equation~\eqref{eq:probability:probability_distributions:dirichlet:pdf} below.

@@

\noindent
The mode of the distribution is given in Equation~\eqref{eq:probability:probability_distributions:bernoulli:mode} below.

@@

\subsection{Binomial probability distribution}\label{sec:probability:probability_distributions:bin}

The \index{Binomial probability distribution}Binomial probability distribution is a discrete probability distribution over a random variable $x$ taking on a number of successes in $N$ sequential independent experiments that each ask a \textit{yes-no} question~\cite{ref:wackerly:2014}. The probability of a single independent experiment yielding a success, is given as $\theta$ and the \index{Binomial probability distribution}Binomial probability distribution is denoted as $Bin(N, \theta)$ with support $x \in \{0, 1, \dots, N\}$.  It should be noted that the \index{Binomial probability distribution}Binomial probability distribution is the extension of the \index{Bernoulli probability distribution}Bernoulli probability distribution over $N$ independent sequential experiments and thus, each experiment also yields some boolean outcome. When $N=1$, the experiment is referred to as a Bernoulli trial, and the distribution is just a \index{Bernoulli probability distribution}Bernoulli probability distribution. When $N > 1$, the sequence of outcomes is referred to as a Bernoulli process. The \acf{PMF} of the \index{Binomial probability distribution}Binomial probability distribution is given in Equation~\eqref{eq:probability:probability_distributions:binomial:pmf} below.

@@

\noindent
Similar to the \index{Bernoulli probability distribution}Bernoulli probability distribution, the mean of the \index{Binomial probability distribution}Binomial probability distribution is $N\theta$ given the \acs{CLT} as was shown in Figure \ref{fig:probability:probability_distributions:bernoulli:coin}. The expected value of the \index{Binomial probability distribution}Binomial probability distribution is thus given in Equation~\eqref{eq:probability:probability_distributions:binomial:expected_value} below.

@@

The \index{Categorical probability distribution}Categorical probability distribution is a discrete probability distribution over some random variable $x$, taking on any one of $K$ possible categories~\cite{ref:wackerly:2014}. There is no ordering to these categories and therefore, for simplicity, each category is assigned a numerical representative value such that $k = (1, 2, \dots, K)$. The probabilities for all outcomes is given by the probability vector $\vec{\theta} = (\theta_{1}, \dots, \theta_{K})$.  This means that the probability $P(x=k)=\theta_{k}$, has support $x \in \{1, \dots, K\}$. The \index{Categorical probability distribution}Categorical probability distribution, denoted $Cat(\theta)$, is the generalization of the \index{Bernoulli probability distribution}Bernoulli probability distribution and is sometimes referred to it by its alternative names, the generalised \index{Bernoulli probability distribution}Bernoulli probability distribution or the Multinoulli probability distribution. In probability theory, the \index{Categorical probability distribution}Categorical probability distribution is often used to explain the outcome of a single experiment with more than two possible outcomes, such as rolling a six-sided die~\cite{ref:wackerly:2014}. The \acs{PMF} of the \index{Categorical probability distribution}Categorical probability distribution is given in Equation~\eqref{eq:probability:probability_distributions:categorical:pmf} below.

@@

The \index{Multinomial probability distribution}Multinomial probability distribution is a discrete probability distribution over some random variable $x = (x_{1}, \dots\, x_{K})$ that takes on the counts for each occurrence of $K$ possible classes in $N$ independent trials~\cite{ref:wackerly:2014}. The probabilities for all possible outcomes in a single trial is given by the probability vector $\vec{\theta} = (\theta_{1}, \dots, \theta_{K})$. The \index{Multinomial probability distribution}Multinomial probability distribution, denoted $Mul(N, K, \theta)$, is thus the generalization of the \index{Binomial probability distribution}Binomial probability distribution to $K$ dimensions. Consider the following special cases.

\begin{itemize}
      \item When $K$ is 2 and $N = 1$, the \index{Multinomial probability distribution}Multinomial probability distribution is the \index{Bernoulli probability distribution}Bernoulli probability distribution.
      \item When $K$ is 2 and $N > 1$, the \index{Multinomial probability distribution}Multinomial probability distribution is the \index{Binomial probability distribution}Binomial probability distribution.
      \item When $K > 2$ and $N = 1$, the \index{Multinomial probability distribution}Multinomial probability distribution is the \index{Categorical probability distribution}Categorical probability distribution.
\end{itemize}

@@

\noindent
Similar to the \index{Categorical probability distribution}Categorical probability distribution, the random variable $x$ can also be encoded in binary format, yielding an $N \times K$ matrix $X$ of \index{Bernoulli probability distribution}Bernoulli probability distributions. The support is then given as $X \in \{0, 1\}^{N \times K}, \forall_{i=1}^{N}\sum_{k=1}^{K} x_{i,k} = 1$ so that the standard $K-1$ probability simplex $S$ still holds for each trial. The \acs{PMF} of the \index{Multinomial probability distribution}Multinomial probability distribution can then be rewritten as is presented in Equation~\eqref{eq:probability:probability_distributions:multinomial:pmf_alt} below.

@@

\noindent
In Equation~\eqref{eq:probability:probability_distributions:multinomial:pmf_alt} above, $N_{k}$ is a summary variable, denoting the number of times a category $k$ occurs over all trials in $N$.

@@

\citeauthor{ref:wackerly:2014}\cite{ref:wackerly:2014} states that conjugate priors are prior probability distributions that results in posterior distributions that are of the same functional form $\mathcal{A}(v)$ as the prior, but with different parameter values. This section considers the conjugate priors that are used with the Binomial likelihood and Categorical/Multinomial likelihood.

@@

      \item Let $X = (x_{1}, \dots, x_{I})$ be the outcomes of independent, identical random events, each with boolean outcome. That is $x_{i} | \theta \overset{\text{iid}}{\sim} Ber(\theta)$ and $\mathcal{L}(x_{i} \vert \theta)$ is the Bernoulli likelihood.

@@

\noindent
The Bernoulli likelihood is given in Equation~\eqref{eq:probability:conjugate_priors:binom_likelihood:likelihood} below.

@@

\noindent
The posterior distribution has the same functional form $\mathcal{A}(v)$ as the prior, but with updated prior parameters $\alpha' = N_{1} + \alpha$ and $\beta' = N_{0} + \beta$. This shows that the \index{Beta probability distribution}Beta probability distribution is the conjugate prior used with the Bernoulli likelihood.

@@

The conjugate prior to a \index{Categorical probability distribution}Categorical and \index{Multinomial probability distribution}Multinomial probability distribution is the \index{Dirichlet probability distribution}Dirichlet probability distribution~\cite{ref:wackerly:2014}. The proof of the conjugate prior for the \index{Bernoulli probability distribution}Bernoulli probability distribution is similar to the proof presented in Section \ref{sec:probability:conjugate_priors:binom_likelihood} above. This means that the posterior distribution must have the same functional form $\mathcal{A}(v)$ as the prior distribution. The proof is shown as follows. \\\\
\textbf{Setup}:

@@

      \item Let $X = (x_{1}, \dots, x_{I})$ be the outcomes of independent, identical random events, each with $K$ possible outcomes. That is $x_{i} | \theta \overset{\text{iid}}{\sim} Cat(\theta)$ and $\mathcal{L}(x_{i} \vert \theta)$ is the Categorical likelihood.

@@

\noindent
The likelihood of the Categorical and \index{Multinomial probability distribution}Multinomial probability distributions is given in Equation~\eqref{eq:probability:conjugate_priors:mult_likelihood:likelihood} below.

@@

\noindent
The posterior distribution has the same form $\mathcal{A}(v)$ as the prior, but with updated prior parameters $\alpha' = N + \alpha$. This shows that the \index{Dirichlet probability distribution}Dirichlet probability distribution is the conjugate prior used with the Categorical/Multinomial likelihood.

@@

In general, there are two main views to probability and statistics. These include the \textit{frequentist} and the \textit{Bayesian} view of statistics. Naturally, Bayesian statistics is based on Bayes' Theorem as was presented in Section \ref{sec:probability:bayes_theorem}. Bayesian statistics describe the probability of an event in terms of some belief, based on previous knowledge of the event, and the conditions under which the event happens~\cite{ref:hackenberger:2019}. To introduce the concept of Bayesian inference and analysis, one first need to understand the difference between these two approaches. Bayesian statistics out-date the frequentist approach, but lacked interest in the early days, partly because of the limited applications where the conjugate priors where known~\cite{ref:hackenberger:2019}. More recent advancements in mathematical methods popularised the Bayesian approach again. A notable contribution to this switch, was the development of \acf{MCMC} in the 1950's. This family of algorithms allowed for the construction of random sampling algorithms from a probability distribution, which allows for the calculation of Bayesian hierarchical models~\cite{ref:hackenberger:2019}. Soon after followed one of the earliest papers that use Bayesian statistics in the field of medicine in 1982~\cite{ref:ashby:2006}.

The difference between the frequentist approach and the Bayesian approach can be illustrated using an example.~\citeauthor{ref:hackenberger:2019}~\cite{ref:hackenberger:2019} suggests an experiment that investigates whether the sex ration in some hypothetical mice population is $1:1$. Two experiments can be designed. In the first experiment, mice are randomly selected until the first male is chosen. The result in this experiment will then be the total number of mice chosen by sex. For the second experiment, exactly seven mice are randomly selected. The result of the second experiment would be the number of males and females in a sample of seven. Suppose the outcome of the experiment was $FFFFFFM$, where $F$ represents a female and $M$ represents a male. If the experimental design is not known ahead of time, the result is useless. Consider the $P$-value for each of these experiments. The P-value is the probability of obtaining results at least as extreme as the observed results of a statistical hypothesis test~\cite{ref:beers:2022}. For the first experiment, the $P$-value is $0.031$ and for the second experiment, the $P$-value is $0.227$. Using a confidence level of $\alpha = 0.05$, one would conclude opposite outcomes for these two experiments, when it comes to rejecting the null hypothesis, despite using the same data. The reason for the difference in outcomes, is due to the difference in their null distributions. The first approach uses a geometrical approach, and the second used a binomial approach as is illustrated on Figure \ref{fig:probability:bayesian_statistics:mouse_experiment_outcome} below.

\begin{figure}[htbp]
      \centering
      \includegraphics[width=0.6\textwidth]{images/mouse_experiment_outcome.pdf}
      \caption{The experimental outcomes for the mice-population experiments as was taken from~\cite{ref:hackenberger:2019} .}
      \label{fig:probability:bayesian_statistics:mouse_experiment_outcome}
\end{figure}

\noindent
If Bayesian statistics is used, the experimental design that was chosen does not matter. In Bayesian statistics it is common to use a \index{Beta probability distribution}Beta probability distribution as a prior distribution. If the prior distribution is sampled from $Beta(3,3)$ then using Bayesian analysis, the posterior distribution, according to the outcomes of this experiment, would yield $Beta(9,4)$.

~\citeauthor{ref:hackenberger:2019}~\cite{ref:hackenberger:2019} mentions that the $Beta$ distribution can be seen as a probability distribution of the occurrence of specific parameters. From the information that is now known about the $Beta$ distribution, it is possible to calculate the probability that the sex ratio in this mice population is not $1:1$, with the $Beta$ distribution as a prior, yielding a $P$-value of $0.92$. This means that there is a probability of $92\%$ that the mice population is not $1:1$, regardless of experiment design. An illustration of this is given in Figure \ref{fig:probability:bayesian_statistics:mouse_distributions} below.

\begin{figure}[htbp]
      \centering
      \includegraphics[width=0.6\textwidth]{images/mouse_experiment_distributions.pdf}
      \caption{An illustration of the prior and posterior probability distributions for the outcomes of the mice-population experiment, using a $Beta$ prior, as was taken from~\cite{ref:hackenberger:2019}.}
      \label{fig:probability:bayesian_statistics:mouse_distributions}
\end{figure}

\noindent
It can be seen from Figure \ref{fig:probability:bayesian_statistics:mouse_distributions} how the posterior distribution differs from the prior distribution. This is the nature of Bayesian analysis and is presented in more detail in the following section.

@@

\noindent
The marginal density or mass function of $Y_{1}, Y_{2}, \dots, Y_{N}$ is given in Equation~\eqref{eq:probability:bayesian_statistic:bayesian_analysis:marginal_density} below.

@@

\noindent
Finally, the posterior density of $\theta \vert y_{1}, y_{2}, \dots, y_{n}$, according to Bayes' Theorem, is given in Equation~\eqref{eq:probability:bayesian_statistic:bayesian_analysis:posterior_density} below.
